Концентрация юрфирм и доступность юруслуг: как они связаны?

### 2024-05-17 legal-companies-concentration-and-availability/journal-article/Article.Rmd
Proofreading of introduction, CSL to Chicago, #2

**Аннотация:** The article examines the relationship between the concentration of legal companies and the availability of legal services. To conduct the analysis, a dataset was created using open data from the Federal Tax Service of Russia. The dataset includes information on the location, revenue, expenditure, and number of employees of small and medium-sized legal companies. The concentration is calculated by dividing the number of employees of legal companies in the city by the total number of employees of all legal companies in Russia. Availability is measured as the number of employees of legal companies in the city per 100,000 population. Data visualization and linear regression were utilized to estimate the hypothesized relationship. The study showed a positive linear relationship between concentration and availability. However, there are multiple approaches to achieving good availability of legal services.

**Кілт сөздер:** legal services, concentration, availability, small and medium-sized businesses, open data

@@

**Аннотация:** The article examines the relationship between the concentration of legal companies and the availability of legal services. To conduct the analysis, a dataset was created using open data from the Federal Tax Service of Russia. The dataset includes information on the location, revenue, expenditure, and number of employees of small and medium-sized legal companies. The concentration is calculated by dividing the number of employees of legal companies in the city by the total number of employees of all legal companies in Russia. Availability is measured as the number of employees of legal companies in the city per 100,000 population. Data visualization and linear regression were utilized to estimate the hypothesized relationship. The study showed a positive linear relationship between concentration and availability. However, there are multiple approaches to achieving good availability of legal services.

@@

Geography of services is an important and well-established field of geographical studies. Its history dates back to the second half of 20th century [@Tkachenko2016]. Two research approaches exist in geography of services: the first can be described as economic, and the second can be designated as social. The economic approach, as its name suggests, treats the services as a regular part of the economy, dealing with their role in GDP or employment, their impact on the growth, the interplay between services and other business activities. For example, a research by D. Wójcik [-@Wojcik2020] provides an overview of FihTech geography in an economic-based manner. The social approach, while closely related to the economic one, focuses on the services as the foundation of people's well-being and studies services provision, availability and accessibility, as well as the diversity and quaility of services. For instance, the study in Indonesia [@Pratiwi2021] uses the various socio-economic data from the national health insurance system grouped by ~500 country's districts to analyze the relationship between a set of public health quality metrics and discuss, among the other problems, the equity problems arising from the geographical differences. The issue of access to various services has been of great importance for decades [@Blacksell1990; @Keleher1996] because it is directly connected to the quality of life and ensuring of human rights. Research has also been conducted on the links between market properties and service availability [@Lu2020].

Cities have traditionally been viewed as regional centers that provide goods and services to smaller surrounding settlements [@Christaller1966]. The urban studies are thus tightly connected with the geography of services [@Reilly1931; @Brush1953]. It is well-known that professional and business services tend to cluster in large cities, and their spatial distribution is closely linked to urban hierarchies [@Keeble2002; @Rubalcaba2013; @Kekezi2019]. The concentration of businesses provides economic benefits for the companies and thus is likely to become self-sustaining. The common sense suggests that such concentration does not only leads to the economic growth, but also positively impacts the service provision, thus being productive in both economic and social dimensions. It is usually the largest cities or the regional centres where the local communities benefit from the variety and quantity of diverse service providers, compared to smaller settlements and rural areas, where the availability of services is limited due to the low number of their providers. However, is this relationship true indeed? The common sense can easily turn out to be wrong, and thus the rigorous data-driven testing of the connection between the number of service providers and the provision of services is crucial.

This paper empirically examines the relationship between the number of law firms in Russian cities and the provision of legal services. Typically, the literature on services availability focuses on medical services [@Joseph1982], but seems that legal services, though being understudied by geographers, are also worth attention [@Blacksell1988; @Patel2008], because they are directly related to the rule of law, protection of human rights, supporting orderliness and stability of social relations. The analysis is based on the open data about small and medium-sized businesses (SMB) published by the Federal Tax Service (FTS) of Russia. The main research question is how the specific availability of legal professionals is affected by the concentration of law firms. Additionally, the study aims to find out geographical patterns in the spatial distribution of legal services availability.

>>>>

**Аннотация:**

**Кілт сөздер:** заң қызметтері, заң фирмалары, қауіпсіздік, қалалар, нарықтың қанықтығы, шағын және орта бизнес, ашық деректер

@@

**Аннотация:**

@@

Geography of services is a significant and well-established branch of geographical studies. The discipline's history can be traced back to the second half of the 20th century, as documented by @Tkachenko2016. Two distinct research approaches exist within the field of geography of services. The first may be described as economic, while the second may be designated as social. The economic approach, as its name suggests, treats the services as a regular part of the economy. It considers their role in GDP or employment, their impact on growth, and the interplay between services and other business activities. For instance, a study by @Wojcik2020 offers an economic-based overview of FihTech geography. The social approach, while closely related to the economic one, focuses on the services as the foundation of people's well-being and studies the provision, availability and accessibility of services, as well as their diversity and quality. For example, the study in Indonesia [@Pratiwi2021] employs a variety of socio-economic data from the national health insurance system to discuss, among other issues, the equity problems that arise from geographical differences. A problem of access to various services has been a significant concern for decades [@Blacksell1990; @Keleher1996], as it is directly related to the quality of life and the guarantee of human rights. Furthermore, research has been conducted on the relationship between geographical properties and service availability [@Holzer2000; @Johnson2006].

Cities have traditionally been regarded as regional hubs that provide goods and services to smaller surrounding settlements [@Christaller1966]. Consequently, urban studies are closely intertwined with the geography of services [@Reilly1931; @Brush1953]. It is widely acknowledged that professional and business services tend to cluster, and their spatial distribution is closely linked to urban hierarchies [@Bennett1999; @Keeble2002; @Rubalcaba2013; @Kekezi2019]. The concentration of businesses provides economic benefits for the companies and, as such, the concentration is likely to become self-sustaining. It is reasonable to assume that such concentration does not only leads to the economic growth, but also positively impacts the service provision, thereby being productive in both economic and social dimensions. It is a common observation that the largest cities and regional centers offer the greater variety and quantity of diverse service providers than smaller settlements and rural areas, where the availability of services is limited due to the low number of their providers. Nevertheless, it must be asked: is this relationship truly valid? The common-sense notion that the number of service providers is correlated with the provision of services may, in fact, be incorrect. Thus, rigorous data-driven testing of this hypothesis is necessary.

This paper empirically examines the relationship between the number of law firms in Russian cities and the provision of legal services. The existing literature on the availability of services typically focuses on medical services [@Joseph1982], with relatively little attention paid to legal services. Nevertheless, there is a growing recognition that spatial aspects of legal services, though understudied by geographers, are worthy of investigation [@Blacksell1988; @Patel2008]. This is because they are directly related to the rule of law, the protection of human rights, and the support of social order and stability. The analysis is based on the open data about small and medium-sized businesses (SMB) published by the Federal Tax Service (FTS) of Russia. The primary research question is how the specific provision of legal professionals is influenced by the number of lawyers. Additionally, the study aims to identify geographical patterns in the spatial distribution of legal services provision.
